---
title: "IE6200 Final_Hung"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
##read.csv("/cloud/project/SuicideData.csv")
data <- read.csv("SuicideData.csv")
head(data)
summary(data)
```

```{r}
##Subset data where country = United States
data.US <- subset(data, data$country=="United States")
##Calculate the total suicide each year from all age ranges
us1985 <- sum(data.US$suicides_no[data.US$year=="1985"])
us1986 <- sum(data.US$suicides_no[data.US$year=="1986"])
us1987 <- sum(data.US$suicides_no[data.US$year=="1987"])
us1988 <- sum(data.US$suicides_no[data.US$year=="1988"])
us1989 <- sum(data.US$suicides_no[data.US$year=="1989"])
us1990 <- sum(data.US$suicides_no[data.US$year=="1990"])
us1991 <- sum(data.US$suicides_no[data.US$year=="1991"])
us1992 <- sum(data.US$suicides_no[data.US$year=="1992"])
us1993 <- sum(data.US$suicides_no[data.US$year=="1993"])
us1994 <- sum(data.US$suicides_no[data.US$year=="1994"])
us1995 <- sum(data.US$suicides_no[data.US$year=="1995"])
us1996 <- sum(data.US$suicides_no[data.US$year=="1996"])
us1997 <- sum(data.US$suicides_no[data.US$year=="1997"])
us1998 <- sum(data.US$suicides_no[data.US$year=="1998"])
us1999 <- sum(data.US$suicides_no[data.US$year=="1999"])
us2000 <- sum(data.US$suicides_no[data.US$year=="2000"])
us2001 <- sum(data.US$suicides_no[data.US$year=="2001"])
us2002 <- sum(data.US$suicides_no[data.US$year=="2002"])
us2003 <- sum(data.US$suicides_no[data.US$year=="2003"])
us2004 <- sum(data.US$suicides_no[data.US$year=="2004"])
us2005 <- sum(data.US$suicides_no[data.US$year=="2005"])
us2006 <- sum(data.US$suicides_no[data.US$year=="2006"])
us2007 <- sum(data.US$suicides_no[data.US$year=="2007"])
us2008 <- sum(data.US$suicides_no[data.US$year=="2008"])
us2009 <- sum(data.US$suicides_no[data.US$year=="2009"])
us2010 <- sum(data.US$suicides_no[data.US$year=="2010"])
us2011 <- sum(data.US$suicides_no[data.US$year=="2011"])
us2012 <- sum(data.US$suicides_no[data.US$year=="2012"])
us2013 <- sum(data.US$suicides_no[data.US$year=="2013"])
us2014 <- sum(data.US$suicides_no[data.US$year=="2014"])
us2015 <- sum(data.US$suicides_no[data.US$year=="2015"])
```

```{r}
##Create a data frame to plot QQ plot
UStotal <- data.frame("Year" = 1985:2015, TotalSuicide = c(us1985, us1986, us1987, us1988, us1989, us1990, us1991, us1992, us1993, us1994, us1995, us1996, us1997, us1998, us1999, us2000, us2001, us2002, us2003, us2004, us2005, us2006, us2007, us2008, us2009, us2010, us2011, us2012, us2013, us2014, us2015))
UStotal
```

### ONE SAMPLE t-TEST - Traditional

Null Hypothesis: mu = 47085 (The true mean suicide number per year is 47085)
Alternate Hypothesis: mu  != 47085 (The true mean suicide number per year is different than 47085)

```{r}
##QQ plot
qqnorm(UStotal$TotalSuicide)
```

```{r}
##sample mean
x_bar1 <- mean(UStotal$TotalSuicide)
##mu_0
mu_01 <- 47085
##sample SD
s1 <- sd(UStotal$TotalSuicide)
##number of years
n1 <- 31
##t-test test statistic
t1 <- (x_bar1 - mu_01)/(s1/sqrt(n1))
##two-sided p-value
two_sided_t_p1 <- pt(q = t1, df = n1-1)*2
two_sided_t_p1
```

```{r}
##CI
lower1 <- x_bar1 + (qt(0.025, n1-1)*(s1/sqrt(n1)))
upper1 <- x_bar1 + (qt(0.975, n1-1)*(s1/sqrt(n1)))
lower1
upper1
```

```{r}
t.test(UStotal$TotalSuicide, alternative = "two.sided", mu = 47085)
```

### ONE SAMPLE t-Test - Bootstrap Approach
```{r}
##install.packages("funModeling")
##install.packages("Hmisc")
library("Hmisc")
library("funModeling")
profiling_num(UStotal)
```


```{r}
##Simulations 100 times
set.seed(111)
num_sims1 <- 10000
##A vector to store my results
results1 <- rep(NA, num_sims1)
##A loop for completing the simulation
for(i in 1:num_sims1){
  results1[i] <- mean(sample(x = UStotal$TotalSuicide, size = n1, replace = TRUE))
}
##Finally plot the results
hist(results1, freq = FALSE, main = "Sampling Distribution of the Sample Mean", xlab = "Average Suicide Numbers", ylab = "Density")
##Estimate a normal curve over it
lines(x = seq(31000,36000, 1), dnorm(seq(31000,36000, 1), mean = x_bar1, sd = sd(UStotal$TotalSuicide)/sqrt(n1)))
```

```{r}
##Shift the sample so that the null hypothesis is true
set.seed(111)
time_given_H0_true1 <- UStotal$TotalSuicide - mean(UStotal$TotalSuicide) + mu_01
##Simulations
num_sims2 <- 10000
##A Vector to store my results
results_given_h0_true1 <- rep(NA, num_sims2)
##A loop for completing the simulation
for(i in 1:num_sims2){
  results_given_h0_true1[i] <- mean(sample(x = time_given_H0_true1, size = n1, replace = TRUE))
}
#Finally Plot the results
hist(results_given_h0_true1, freq = FALSE, main = "Sampling Distribution Null is True", xlab = "Number of Suicide", ylab = "Density")

```

```{r}
set.seed(111)
##More extreme on lower end
low_end_extreme <- mean(results_given_h0_true1)+(mean(results_given_h0_true1)-x_bar1)
##Counts of values more extreme than the test statistic in our original sample, given H0 is true
##Two sided given the alternate hypothesis
count_of_more_extreme_lower <- sum(results_given_h0_true1 >= low_end_extreme)
count_of_more_extreme_upper <- sum(results_given_h0_true1 <= x_bar1)

bootsrtap_p1 <- (count_of_more_extreme_lower + count_of_more_extreme_upper)/num_sims2
bootsrtap_p1
two_sided_t_p1
```

```{r}
##CI of bootstrap
##Need the standard error which is the sd of the results
bootstrap_SE_x_bar1 <- sd(results1)
##An estimate is to use the formula statisitc +/- 2*SE
c(x_bar1 - 2*bootstrap_SE_x_bar1, x_bar1 + 2*bootstrap_SE_x_bar1)
```

```{r}
##Find 5th and 95th quantiles to determine the bounds
c(quantile(results1, c(0.025, 0.975)))
```

```{r}
##Compare t-methods
c(x_bar1 + (qt(0.025, n1-1)*(s1/sqrt(n1))), x_bar1+qt(0.975, n1-1)*(s1/sqrt(n1)))
```

### ONE SAMPLE PROPORTION TEST - Traditional

```{r}
totalUS_R1 <- sum(data.US$suicides_no[data.US$age=="5-14 years"])
totalUS_R2 <- sum(data.US$suicides_no[data.US$age=="15-24 years"])
totalUS_R3 <- sum(data.US$suicides_no[data.US$age=="25-34 years"])
totalUS_R4 <- sum(data.US$suicides_no[data.US$age=="35-54 years"])
totalUS_R5 <- sum(data.US$suicides_no[data.US$age=="55-74 years"])
totalUS_R6 <- sum(data.US$suicides_no[data.US$age=="75+ years"])
totalUS <- sum(data.US$suicides_no)
```

```{r}
z1 <- ((totalUS_R3/totalUS)-(1/6)) / sqrt(((1/6)*(5/6))/totalUS)
z1
```

```{r}
binom.test(x = totalUS_R3, n = totalUS, p=(1/6), alternative = "greater")
```

```{r}
pnorm(z1, lower.tail = FALSE)
```

```{r}
binom.test(x = totalUS_R3, n = totalUS, p=(1/6), alternative = "greater")$conf.int
```

```{r}
##normal approx - greater
c((totalUS_R3/totalUS) - 1.64*(((totalUS_R3/totalUS)*(1-(totalUS_R3/totalUS)))/totalUS),1)
```

```{r}
binom.test(x = totalUS_R3, n = totalUS, p=(1/6), alternative = "two.sided")$conf.int
```

```{r}
##normal approx - two.sided
c((totalUS_R3/totalUS) - 1.64*(((totalUS_R3/totalUS)*(1-(totalUS_R3/totalUS)))/totalUS),(totalUS_R3/totalUS) + 1.64*(((totalUS_R3/totalUS)*(1-(totalUS_R3/totalUS)))/totalUS))
```

```{r}
## One proportion for 101 countries
total_R1 <- sum(data$suicides_no[data$age=="5-14 years"])
total_R2 <- sum(data$suicides_no[data$age=="15-24 years"])
total_R3 <- sum(data$suicides_no[data$age=="25-34 years"])
total_R4 <- sum(data$suicides_no[data$age=="35-54 years"])
total_R5 <- sum(data$suicides_no[data$age=="55-74 years"])
total_R6 <- sum(data$suicides_no[data$age=="75+ years"])
total <- sum(data$suicides_no)
```

```{r}
##binom test for 101 countries - one sided
binom.test(x = total_R3, n = total, p=(1/6), alternative = "greater")
```

```{r}
##binom test for 101 countries - two sided
binom.test(x = total_R3, n = total, p=(1/6), alternative = "two.sided")

## NO CI Yet
```

### ONE SAMPLE TEST of PROPORTION - Boostrap Approach

```{r}
Range3 <- factor(rep(c("25-34 years", "other"), c(totalUS_R3, (totalUS - totalUS_R3))))
table(Range3)
```

```{r}
##Easier to use for bootstrapping
Range3_ <- rep(c(1, 0), c(totalUS_R3, (totalUS - totalUS_R3)))
```

```{r}
set.seed(111)
##Simulations
num_sims3 <- 1000
##A vector to store results
results2 <- rep(NA, num_sims3)
##A loop for completing the simulations
for(i in 1:num_sims3){
  results2[i] <- mean(sample(x = Range3_, size = totalUS, replace = TRUE))
}
##Finally plot
hist(results2, freq = FALSE, main = "Sampling Distribution of the Sample Proportion", xlab = "Proportion of 25-34 years old", ylab = "Density")
##Estimate a normal curve over it
lines(x = seq(0.1745, 0.1775, 0.0001), dnorm(seq(0.1745, 0.1775, 0.0001), mean = mean(results2), sd = sd(results2)))
```

```{r}
##CI of bootstrap
c(quantile(results2, c(0.05, 1)))
```

```{r}
binom.test(x = totalUS_R3, n = totalUS, p=(1/6), alternative = "greater")$conf.int
```

```{r}
c((totalUS_R3/totalUS) - 1.64*sqrt(((totalUS_R3/totalUS)*(1-(totalUS_R3/totalUS)))/totalUS),1)
```

```{r}
Range3_2 <- rep(c(1, 0), c(totalUS/6, (totalUS*5/6)))
```

```{r}
##Under Null is TRUE
set.seed(111)
##Simulations
num_sims4 <- 100
##A vector to store results
results3 <- rep(NA, num_sims4)
##A loop for completing the simulations
for(i in 1:num_sims4){
  results3[i] <- mean(sample(x = Range3_2, size = totalUS, replace = TRUE))
}
##Finally plot
hist(results3, freq = FALSE, main = "Sampling Distribution of the Sample Proportion", xlab = "Proportion of 25-34 years old", ylab = "Density")
##Estimate a normal curve over it
lines(x = seq(0.1655, 0.1690, 0.0001), dnorm(seq(0.1655, 0.1690, 0.0001), mean = mean(results3), sd = sd(results3)))
abline(v = (1/6), col="red")
```

```{r}
propR3 <- totalUS_R3/totalUS
##Count extreme
count_of_more_extreme_uppertail <- sum(results3 >= propR3)
bootstrap_p2 <- count_of_more_extreme_uppertail/num_sims4
bootstrap_p2
```

```{r}
##Exact Binomial p-value
binom.test(x = totalUS_R3, n = totalUS, p=(1/6), alternative = "greater")$p.value
```

```{r}
##Normal Approx
pnorm(z1, lower.tail = FALSE)
```

### TWO SAMPLE t-Test - Traditional

```{r}
##two populations are female and male, with numbers of suicide per 100k people
##install.packages("tidyverse")
##library(tidyverse)
##data.US  %>% select(2:7)
```

```{r}
##QQ plot of suicide number of male and female in different age ranges
qqnorm(data.US$suicides.100k.pop)
```

```{r}
##QQ Plot of suicide number of female in different age ranges
qqnorm(data.US$suicides.100k.pop[data.US$sex=="female"])
```

```{r}
##QQ Plot of suicide number of male in different age ranges
qqnorm(data.US$suicides.100k.pop[data.US$sex=="male"])
```


```{r}
##sample means
x_bar_f <- mean(data.US$suicides.100k.pop[data.US$sex=="female"])
x_bar_m <- mean(data.US$suicides.100k.pop[data.US$sex=="male"])
##null hypothesized population mean difference between the two groups
mu_02 <- 0
##sample variances
s_f_sq <- sd(data.US$suicides.100k.pop[data.US$sex=="female"])**2
s_m_sq <- sd(data.US$suicides.100k.pop[data.US$sex=="male"])**2
##sample size
n_f <- length(data.US$suicides.100k.pop[data.US$sex=="female"])
n_m <- length(data.US$suicides.100k.pop[data.US$sex=="male"])
##t-test test statistic
t2 <- (x_bar_f - x_bar_m - mu_02)/sqrt((s_f_sq/n_f)+(s_m_sq/n_m))
##one sided upper p-value
one_sided_diff_t_p <- pt(q=t2, df=(min(n_f, n_m)-1), lower.tail = TRUE)*2
one_sided_diff_t_p
```

```{r}
##lower bound
(x_bar_f - x_bar_m) + (qt(0.025, min(n_f, n_m)-1)*sqrt((s_f_sq/n_f)+(s_m_sq/n_m)))
```

```{r}
##upper bound
(x_bar_f - x_bar_m) + (qt(0.975, min(n_f, n_m)-1)*sqrt((s_f_sq/n_f)+(s_m_sq/n_m)))
```

```{r}
##use t-test function in R
t.test(data.US$suicides.100k.pop[data.US$sex=="female"],data.US$suicides.100k.pop[data.US$sex=="male"])
```

### TWO SAMPLE Diff in Mean - Bootstrap Approach

```{r}
##install.packages('mosaic')
require(mosaic)
set.seed(111)
x_hist1 <- data.US %>%
  filter(data.US$sex== "female")
x_hist1 <- do(1000) * mean(sample(data.US$suicides.100k.pop,20))
h1 <- hist(x_hist1$mean, main = "Sampling Distribution For Female", xlab = "Mean of Suicides per 100k people", prob = T, col = "grey")
lines(density(x_hist1$mean), col = "blue", lwd = 2)
h1

require(mosaic)
set.seed(111)
x_hist0 <- data.US %>%
  filter(data.US$sex=="male")
x_hist0 <- do(1000) * mean(sample(data.US$suicides.100k.pop,20))
h0 <- hist(x_hist1$mean, main = "Sampling Distribution For Male", xlab = "Mean of Suicides per 100k people", prob = T, col = "grey")
lines(density(x_hist0$mean), col = "blue", lwd = 2)
h0
```

```{r}
num_sims5 <- 1000
##A vector to store results
results4 <- rep(NA, num_sims5)
##A loop for completing the simulations
for(i in 1:num_sims5){
  mean_female <- mean(sample(x = data.US$suicides.100k.pop[data.US$sex=="female"], size = n_f, replace = TRUE))
  mean_male <- mean(sample(x = data.US$suicides.100k.pop[data.US$sex=="male"], size = n_m, replace = TRUE))
  results4[i] <- mean_female - mean_male
}
##Finally plot
hist(results4, freq = FALSE, main = "Sampling Distribution of the Sample Mean", xlab = "Number of Suicide Difference", ylab = "Density")
```

```{r}
##Bootstrap one-sided CI
c(quantile(results4, c(0.025, 0.975)))
```

```{r}
##T-method
t.test(data.US$suicides.100k.pop[data.US$sex=="female"],data.US$suicides.100k.pop[data.US$sex=="male"])$conf.int
```

```{r}
set.seed(111)
num_sims6 <- 1000
##A vector to store results
results_given_h0_true2 <- rep(NA, num_sims6)
##A loop for completing the simulation
for(i in 1:num_sims6){
  shuffled_groups <- transform(data.US, sex = sample(sex))
  mean_female2 <- mean(shuffled_groups$suicides.100k.pop[shuffled_groups$sex=="female"])
  mean_male2 <- mean(shuffled_groups$suicides.100k.pop[shuffled_groups$sex=="male"])
  results_given_h0_true2[i] <- mean_female2 - mean_male2
}
##Finally plot
hist(results_given_h0_true2, freq = FALSE, main = "Dist. of the Diff in  Sample Means Under Null", xlab  =  "Average Number of Suicide Difference under Null", ylab = "Density")
diff_in_samplemean <- mean(data.US$suicides.100k.pop[data.US$sex=="female"]) - mean(data.US$suicides.100k.pop[data.US$sex=="male"])
diff_in_samplemean
##abline(v=diff_in_samplemean, col = "blue")
##abline(v=abs(diff_in_samplemean), col = "red")
```
```{r}
##Count of values more extreme
##Two sided given
count_of_more_extreme_lower2 <- sum(results_given_h0_true2 <= diff_in_samplemean)
count_of_more_extreme_upper2 <- sum(results_given_h0_true2 >= abs(diff_in_samplemean))

boostrap_p2 <- (count_of_more_extreme_lower2 + count_of_more_extreme_upper2)/num_sims6
boostrap_p2
```

```{r}
t.test(data.US$suicides.100k.pop[data.US$sex=="female"],data.US$suicides.100k.pop[data.US$sex=="male"])$p.value
```

### TWO SAMPLE DIFF IN PROPORTION - Traditional
```{r}
##suicides in two different popultaions: Female and Male
dataUS_male <- subset(data.US, data.US$sex=="male")
dataUS_female <- subset(data.US, data.US$sex=="female")
```

```{r}
data2 <- aggregate(suicides.100k.pop ~ year+sex, data=data.US, FUN=sum) 
```

```{r}
data3 <- aggregate(suicides_no~ year, data=dataUS_female, FUN=sum)
data4 <- aggregate(population~ year, data=dataUS_female, FUN=sum) 
```

```{r}
##sample props
p_hat_f2 <- sum(dataUS_female$suicides_no)/sum(dataUS_female$population)
p_hat_m2 <- sum(dataUS_male$suicides_no)/sum(dataUS_male$population)
#null hypothesized population pro difference between the two group
p_0 <- 0
#sample size
n_f2 <- sum(dataUS_female$population)
n_m2 <- sum(dataUS_male$population)
#sample variances
den_p_f2 <- (p_hat_f2*(1-p_hat_f2))/n_f2
den_p_m2 <- (p_hat_m2*(1-p_hat_m2))/n_m2
#z-test test statistic
z <- (p_hat_m2 - p_hat_f2 - p_0)/sqrt(den_p_f2 + den_p_m2)
#two sided p-value
two_sided_diff_prop_p<- pnorm(q=z, lower.tail = FALSE)*2
two_sided_diff_prop_p
```

```{r}
##lower bound
(p_hat_f2-p_hat_m2)+(qnorm(0.025)*sqrt(den_p_f2+den_p_m2))
```

```{r}
##upper bound
(p_hat_f2-p_hat_m2)+(qnorm(0.975)*sqrt(den_p_f2+den_p_m2))
```

### TWO SAMPLE Diff in PROPORTION - Bootstrap and Randomization Approach
```{r}
f1 <- sum(dataUS_female$suicides_no)
f1
m_1 <- sum(dataUS_male$suicides_no)
ff <- as.integer(sum(dataUS_female$population)/1000)
ff
mx <- as.integer(sum(dataUS_male$population)/1000)
```


```{r}
fake_f <- rep(c(1,0), c(f1, (ff-f1)))
fake_m <- rep(c(1,0), c(m_1, (mx-m_1)))
```

```{r}
#Make the Data
set.seed(111)
num_sims7 <- 1000
##A vector to store results
results5 <- rep(NA, num_sims7)
##A loop for completing the simulation
for(i in 1:num_sims7){
  prop_female <- mean(sample(x=fake_f, size = ff, replace = TRUE))
  prop_male <- mean(sample(x=fake_m, size = mx, replace = TRUE))
  results5[i] <- prop_female - prop_male
}
##Finally Plot
hist(results5, freq=FALSE, main = "Dist. of the Diff in Prop", xlab = "Difference in Prop. of Suicide", ylab = "Density")
```

```{r}
##Boostrap
c(quantile(results5, c(0.025, 0.975)))
```

```{r}
##Normal Approximation
c((p_hat_f2-p_hat_m2)+(qnorm(0.025)*sqrt(den_p_f2+den_p_m2)), (p_hat_f2-p_hat_m2)+(qnorm(0.975)*sqrt(den_p_f2+den_p_m2)))
```

```{r}
ff <- sum(dataUS_female$population)
mm <- sum(dataUS_male$population)
```

```{r}
#Make Data
df_combined <- data.frame("population" = c(fake_f, fake_m), "sex" = rep(c("female", "male"), c(ff, mx)))
summary(df_combined)
```

```{r}
mean(df_combined$population[df_combined$sex=="female"]) == (p_hat_f2*1000)
```

```{r}
mean(df_combined$population[df_combined$sex=="male"]) == (p_hat_m2*1000)
```

```{r}
set.seed(111)
##A vector to store results
results_given_h0_true3 <- rep(NA, num_sims7)
##A loop for completing the simulation
for(i in 1:num_sims7){
  shuffled_groups2 <- transform(df_combined, sex=sample(sex))
  prop_f2 <- mean(shuffled_groups2$population[shuffled_groups2$sex=="female"])
  prop_m2 <- mean(shuffled_groups2$population[shuffled_groups2$sex=="male"])
  results_given_h0_true3[i] <- prop_f2 - prop_m2
}
##Finally plot
hist(results_given_h0_true3, freq = FALSE, main = "Dist. of the Diff in Sample Sample Props Under Null", xlab = "Average Difference in prop. Suicide under Null", ylab = "Density")
```

```{r}
##Count of values more extreme than the test statistic
#two sided given the alternate hypothesis
diff_in_sampleprop <- p_hat_f2 - p_hat_m2

count_of_more_extreme_lower3 <- sum(results_given_h0_true3/1000 <= diff_in_sampleprop)
count_of_more_extreme_upper3 <- sum(results_given_h0_true3/1000 >= -diff_in_sampleprop)

bootsrtap_p3 <- (count_of_more_extreme_lower3+count_of_more_extreme_upper3)/num_sims7

```

### Chi-Square Goodnees of Fit Test

Null Hypothesis: p_r1 = p_r2 = p_r3 = p_r4 = p_r5 = p_r6 = 1/6
The proportion fo each age group is the same and is equal to 1/6
Alternate Hypothesis: some p_i != 1/6
At least one of the proportions is not equal to 1/6

```{r}
##chi-square with 6 different age ranges
totalR1 <- sum(data.US$suicides_no[data.US$age=="5-14 years"])
totalR2 <- sum(data.US$suicides_no[data.US$age=="15-24 years"])
totalR3 <- sum(data.US$suicides_no[data.US$age=="25-34 years"])
totalR4 <- sum(data.US$suicides_no[data.US$age=="35-54 years"])
totalR5 <- sum(data.US$suicides_no[data.US$age=="55-74 years"])
totalR6 <- sum(data.US$suicides_no[data.US$age=="75+ years"])
totalR <- sum(data.US$suicides_no)
avgR <- totalR/6
```

```{r}
suicide_age <- matrix(c(totalR1, totalR2, totalR3, totalR4, totalR5, totalR6), ncol=6, byrow=TRUE)
colnames(suicide_age) <- c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years")
suicide_age
```

```{r}
prop.table(suicide_age)
```

```{r}
##test statistics
chi <- (((totalR1-avgR)^2)/avgR)+(((totalR2-avgR)^2)/avgR)+(((totalR3-avgR)^2)/avgR)+(((totalR4-avgR)^2)/avgR)+(((totalR5-avgR)^2)/avgR)+(((totalR6-avgR)^2)/avgR)
```

```{r}
##p-value
pchisq(chi, df=6-1, lower.tail=FALSE)
```

```{r}
avgR
totalR/6
```


```{r}
sol_under_H0 <- rep(c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"), avgR)
table(sol_under_H0)
```

```{r}
set.seed(111)
num_sims <- 1000
##A vector to store my results
chisq_stats_under_H0 <- rep(NA, num_sims)
##A loop for completing the simulation
for(i in 1:num_sims){
  new_sample <- sample(sol_under_H0, totalR, replace=T)
  chisq_stats_under_H0[i] <- sum(((table(new_sample)-avgR)^2)/avgR)
}
```

```{r}
hist(chisq_stats_under_H0, freq=FALSE, main = "Dist. of the Chi-Square Statistic Under Null", xlab = "Chi-Square stat under Null", ylab = "Density")
abline(v=sum(((suicide_age - avgR)^2)/avgR), col="red")

```

```{r}
sum(chisq_stats_under_H0 >= sum(((suicide_age)-avgR)^2)/avgR)/ num_sims
```


## Chi-square association between Two Categorical Variable

```{r}
##Null: The number of suicide is not associated with gender
##Alternate: The number of suicide is associated with gender
```

```{r}
male_R1 <- sum(dataUS_male$suicides_no[dataUS_male$age=="5-14 years"])
male_R2 <- sum(dataUS_male$suicides_no[dataUS_male$age=="15-24 years"])
male_R3 <- sum(dataUS_male$suicides_no[dataUS_male$age=="25-34 years"])
male_R4 <- sum(dataUS_male$suicides_no[dataUS_male$age=="35-54 years"])
male_R5 <- sum(dataUS_male$suicides_no[dataUS_male$age=="55-74 years"])
male_R6 <- sum(dataUS_male$suicides_no[dataUS_male$age=="75+ years"])
female_R1 <- sum(dataUS_female$suicides_no[dataUS_female$age=="5-14 years"])
female_R2 <- sum(dataUS_female$suicides_no[dataUS_female$age=="15-24 years"])
female_R3 <- sum(dataUS_female$suicides_no[dataUS_female$age=="25-34 years"])
female_R4 <- sum(dataUS_female$suicides_no[dataUS_female$age=="35-54 years"])
female_R5 <- sum(dataUS_female$suicides_no[dataUS_female$age=="55-74 years"])
female_R6 <- sum(dataUS_female$suicides_no[dataUS_female$age=="75+ years"])
f_m_age <- matrix(c(female_R1, male_R1, female_R2, male_R2, female_R3, male_R3, female_R4, male_R4, female_R5, male_R5, female_R6, male_R6), ncol=2, byrow=TRUE)
colnames(f_m_age) <- c("Female", "Male")
rownames(f_m_age) <- c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years")
f_m_age
```

```{r}
ad <- addmargins(f_m_age)
ad
```

```{r}
addmargins((prop.table(f_m_age)))
```

```{r}
n2 <- addmargins(f_m_age)[7,3]
ex_r1_m <- (ad[1,3]*ad[7,2]/n2)
ex_r1_f <- (ad[1,3]*ad[7,1]/n2)
ex_r2_m <- (ad[2,3]*ad[7,2]/n2)
ex_r2_f <- (ad[2,3]*ad[7,1]/n2)
ex_r3_m <- (ad[3,3]*ad[7,2]/n2)
ex_r3_f <- (ad[3,3]*ad[7,1]/n2)
ex_r4_m <- (ad[4,3]*ad[7,2]/n2)
ex_r4_f <- (ad[4,3]*ad[7,1]/n2)
ex_r5_m <- (ad[5,3]*ad[7,2]/n2)
ex_r5_f <- (ad[5,3]*ad[7,1]/n2)
ex_r6_m <- (ad[6,3]*ad[7,2]/n2)
ex_r6_f <- (ad[6,3]*ad[7,1]/n2)

```

```{r}
chi_sq_stat <- sum(
  ((ad[1,1]-ex_r1_f)^2)/ex_r1_f,
  ((ad[2,1]-ex_r2_f)^2)/ex_r2_f,
  ((ad[3,1]-ex_r3_f)^2)/ex_r3_f,
  ((ad[4,1]-ex_r4_f)^2)/ex_r4_f,
  ((ad[5,1]-ex_r5_f)^2)/ex_r5_f,
  ((ad[6,1]-ex_r6_f)^2)/ex_r6_f,
  ((ad[1,2]-ex_r1_m)^2)/ex_r1_m,
  ((ad[2,2]-ex_r2_m)^2)/ex_r2_m,
  ((ad[3,2]-ex_r3_m)^2)/ex_r3_m,
  ((ad[4,2]-ex_r4_m)^2)/ex_r4_m,
  ((ad[5,2]-ex_r5_m)^2)/ex_r5_m,
  ((ad[6,2]-ex_r6_m)^2)/ex_r6_m
)
chi_sq_stat
```

```{r}
pchisq(chi_sq_stat, df = (6-1)*(2-1), lower.tail = FALSE)
```
